\propertitle

### 2019-06-17 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-17T10:37:37+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The context we consider is as follows. We have recorded the spiking
activities of $n$ neurons for a given amount of time. These spikes are
binned into $T$ time bins and binarized to $\set{0,1}$ values in each bin.
Denote by $\ya_{t}$ the number of neurons that fire during time bin $t$,
divided by the total number of neurons $n$; we call this the
\emph{normalized total activity} or just \enquote{total activity} or
\enquote{activity} for short. Obviously
$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
during time bin $t$; if $\ya_{t}=1$, all spiked at some point during time
bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
  bin $t$}.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain regions
\mynote{be more specific}. The neurons that are recorded within the region
are not, and cannot yet be, specifically chosen from among the rest, owing
to several limiting factors -- for example, limitations in the precision
with which electrodes are inserted, or with which neurons are targeted by
viruses, and so on. In fact, the set of recorded neurons may even change
during very long recordings or across experimental trials. We assume that
there's a region, comprising a network of $N$ neurons, from which other
sets of neurons had the same probability of being recorded as the set of
$n$ neurons that have actually been recorded. We call this the
\enquote{full network}. During the recording, the normalized total
activity of these $N$ neurons at $t$ was $\yA_{t}$. We don't know the
values $\yA_{t}$ at each $t$; we only know for certain that
$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
for obvious reasons. For the moment we assume that we know $N$; in \sect***
we discuss the consequences of our lack of precise knowledge about this
number.

@@

Regarding~\ref{item:maxent_step}, we assume you to be familiar with the
maximum-entropy method. In short, it amounts to a pair of prescriptions:
choose the distribution having maximum Shannon entropy among those
satisfying specific convex constraints, such as fixed expectation values; and
judge these expectation values to be equal to measured averages. We add two
remarks about this method that are seldom made in the literature. First,
the \enquote{distribution} given by this method is the zeroth-order
approximation of four different distributions:
\begin{itemize}
\item the \emph{probability} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
\item the \emph{probability} distribution for the value of the total
  activity at a \emph{new} time bin,
\item the most probable \emph{time-frequency} distribution of the total
  activity across the \emph{recorded} bins,
\item the most probable \emph{time-frequency} distribution of the total
  activity in a very long run of \emph{new} time bins.
\end{itemize}
The maximum-entropy distribution is thus an approximate answer to four
conceptually different inferences. Note that although the four
distributions above numerically coincide in this approximation, they differ
in higher-order approximations. Second, the maximum-entropy method based on
the Shannon entropy implicitly makes specific assumptions about the
probability of the long-run frequency distributions
\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
assumptions in \sect***

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-17T10:56:29+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The context we consider is as follows. We have recorded the spiking
activities of $n$ neurons for a given amount of time. Let's call these
neurons the \enquote{sample}. Their spikes are binned into $T$ time bins and
binarized to $\set{0,1}$ values in each bin. Denote by $\ya_{t}$ the number
of neurons that fire during time bin $t$, divided by the total number of
neurons $n$; we call this the \emph{normalized total activity} of the
sample, or just its \enquote{total activity} or \enquote{activity} for
short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$,
no neuron spiked during time bin $t$; if $\ya_{t}=1$, all spiked at some
point during time bin $t$. For brevity, let's say \enquote{at $t$} for
\enquote{during time bin $t$}.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain regions
\mynote{be more specific}. The neurons that are recorded within the region
are not, and cannot yet be, specifically chosen from among the rest, owing
to several limiting factors -- for example, limitations in the precision
with which electrodes are inserted, or with which neurons are targeted by
viruses, and so on. In fact, the set of recorded neurons may even change
during very long recordings or across experimental trials. We assume that
there's a region, comprising a network of $N$ neurons, from which other
sets of neurons had the same probability of being recorded as the set of
$n$ neurons that has actually been recorded. We call this the \enquote{full
  network}\mynote{how about calling it \enquote{the pool}?}. During the
recording, the normalized total activity of these $N$ neurons at $t$ was
$\yA_{t}$. We don't know the values $\yA_{t}$ at each $t$; we only know for
certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the moment we assume that
we know $N$; in \sect*** we discuss the consequences of our lack of precise
knowledge about this number.

@@

Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
maximum-entropy method. In short, it amounts to a pair of prescriptions:
choose the distribution having maximum Shannon entropy among those
satisfying specific convex constraints, such as fixed expectation values;
and judge these expectation values to be equal to measured averages. We add
two remarks about this method that are seldom made in the literature.
First, the \enquote{distribution} given by this method is the zeroth-order
approximation of four different distributions:
\begin{itemize}
\item the \emph{probability} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
\item the \emph{probability} distribution for the value of the total
  activity at a \emph{new} time bin,
\item the most probable \emph{time-frequency} distribution of the total
  activity across the \emph{recorded} bins,
\item the most probable \emph{time-frequency} distribution of the total
  activity in a very long run of \emph{new} time bins.
\end{itemize}
The maximum-entropy distribution is thus an approximate answer to four
conceptually different inferences. Note that although the four
distributions above numerically coincide in this approximation, they differ
in higher-order approximations. We discuss their differences in \sect***.
Second, the maximum-entropy method based on the Shannon entropy implicitly
makes specific assumptions about the probability of the long-run frequency
distributions \citep{jaynes1986d_r1996,portamana2009,portamana2017}. We
discuss these assumptions in \sect***.
